---
title: "HW 5"
author: "Alejandro Sanchez"
date: "10/26/2020"
output: html_document
---


```{r}
load("C:/Users/alejo/Desktop/Econometrics/ecob2000_lecture1/acs2017_ny_data.RData")
attach(acs2017_ny)
use_varb <- (AGE >= 20) & (AGE <= 65) & (LABFORCE == 2) & (WKSWORK2 > 4) & (UHRSWORK >= 35) & (Hispanic == 1) & (female == 0) & ((educ_hs == 1) | (educ_advdeg == 1))
dat_use <- subset(acs2017_ny,use_varb) # 
detach()
attach(dat_use)
```

```{r}
model_1 <- lm((INCWAGE ~ AGE))
require(stargazer)
stargazer(model_1, type = "text")
```
```{R}
NNobs <- length(INCWAGE)
set.seed(12345) 
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)
```
```{R}
plot(INCWAGE ~ jitter(AGE, factor = 2), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)
to_be_predicted1 <- data.frame(AGE = 20:65, female = 0, educ_hs = 0, educ_advdeg = 1)
to_be_predicted1$yhat <- predict(model_1, newdata = to_be_predicted1)
lines(yhat ~ AGE, data = to_be_predicted1)
```
The results of this basic regression shows us that at a 99% confidence level there is a positive correlation between AGE and Income and we can also see this when we plot the points and see the best fit line with a positive slope indicating that based on the limits I put Income does seem to rise as one gets older even up until retirement age. 

To try and improve on this model I decided I wanted to see if Family size had any impact . I also wanted to see what happens when we make it a quadratic by squaring Age. 

```{R}
model_2 <- lm((INCWAGE ~ AGE + I(AGE^2) + FAMSIZE))
require(stargazer)  
stargazer(model_2, type = "text")
```
```{R}
Nobs <- length(INCWAGE)
set.seed(12345) 
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

to_be_predicted2 <- data.frame(AGE = 20:65, female = 0, educ_hs = 0, educ_advdeg = 1, FAMSIZE = 4)
to_be_predicted2$yhat <- predict(model_2, newdata = to_be_predicted2)

plot(INCWAGE ~ jitter(AGE, factor = 2), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)

lines(yhat ~ AGE, data = to_be_predicted2)
```
After running this regression I see that overall my variables still affect income at a level that is statistically significant. We still see a correlation at a 99% confidence level. Where we really see the change or at least in my case more obvious is when we plot the the function and we see that the income no longer jumps significantly as age increases. If anything we see that the overall income at retirement has gone down when comparing to the first graph and the gorwth in income over time as you grow older is more gradual and smooth. 

```{R}
model_3 <- lm((INCWAGE ~ AGE + I(AGE^2) + I(AGE^3) + I(AGE^4) + FAMSIZE))
require(stargazer)  
stargazer(model_3, type = "text")
```
```{R}
Nobs <- length(INCWAGE)
set.seed(12345) 
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

to_be_predicted2 <- data.frame(AGE = 20:65, female = 0, educ_hs = 0, educ_advdeg = 1, FAMSIZE = 4)
to_be_predicted2$yhat <- predict(model_3, newdata = to_be_predicted2)

plot(INCWAGE ~ jitter(AGE, factor = 2), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)

lines(yhat ~ AGE, data = to_be_predicted2)
```
After adding Age^3 and Age^4 I see that when I plot I certainly get a different line. This time around it shows that a Male once he starts nearing the age of 50 his wage starts to flatten out up until you start nearing age 60 at which point we see a sudden spike in wages. This spiked my curiosity and I tried rationalizing several explanations in my head such as family size affects time required to be spent at home and as you grow older some members of your family may go off to start their own families allowing you to possibly work more hours. However, I am aware that this might not be the actual reason. Another observation I had is that my indepedent variables are no longer are no longer statistically significant but the regression itself is. 

```{R}
Poly_Model <- lm(INCWAGE ~ poly(AGE,3, raw=TRUE), data=dat_use)
with(dat_use, plot(AGE, INCWAGE, col="grey", ylim=c(0, max(INCWAGE)), 
                xlab = "Age", ylab = "Income", main="Data - polynomial fit"))
pred = data.frame(AGE,
                  INCWAGE=predict(Poly_Model, newdata=data.frame(AGE)))
with(pred, lines(AGE, INCWAGE, col='red', type='l', cex=10))
cf = coef(Poly_Model)
D1 = cf[2] + 2*cf[3]*pred$AGE + 3*cf[4]*pred$AGE^2
D1roots = (-2*cf[3] + c(-1,1)*sqrt((2*cf[3])^2 - 4*3*cf[4]*cf[2]))/(2*3*cf[4])
D2atD1roots =  2*cf[3] + 6*cf[4]*D1roots
max_x = D1roots[which(D2atD1roots < 0)]
max_y = cf %*% max_x^(0:3)
points(max_x, max_y, pch=16, col="red")
text(max_x, max_y, label=round(max_y,2), adj=c(0.5,-1), cex=0.8)
```
This graph shows us that the peak wage is $69,125.15 for a Hispanic male with an advanced degree and a Family size of 4. This peak wage happens as we start to near 60 years old. We saw this as well on the previous graph and this cements the observation that there was an increase as you started nearing 60 years old. 

```{R}
require(stargazer)  
stargazer(Poly_Model, type = "text")
```

We see in the stargazer summary the variables in my model 3 are not statistically significant. What I think this is telling me is that at a certain point when you try to find correlation between age and income using famsize as well you lose the significance which means we don't have enough evidence to reject that there is correlation the variables independently but as a whole the model fits pretty well since it is statistically significant.  
